---
title: "kolejkeR-vignette"
author: "Paweł Morgen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{kolejkeR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
knitr::opts_chunk$set(fig.width=10, fig.height=6) 
library(ggplot2)
library(dplyr)
library(rlang)
library(tidyr)
library(hms)
library(ggrepel)
library(pals)
library(kolejkeR)
data("warsaw_queues")
data <-warsaw_queues



dates <- data %>% select(date) %>% group_by(date) %>% summarise(n())

data[['date_time_posix']]<- as.POSIXct(strptime(paste(data[['time']], data[['date']]), format="%H:%M %Y-%m-%d", tz="Europe/Warsaw"))
data[['time_posix']] <- as.POSIXct(strptime(data[['time']], format="%H:%M", tz="Europe/Warsaw"))

parse_served_number <- function(queuer_token_col) {
  as.numeric(gsub("[[:alpha:]](\\d*)", "\\1", queuer_token_col, perl=TRUE))
}

data_with_queuers_count <- data %>%
  mutate(served_people=replace_na(parse_served_number(aktualnyNumer),0)) %>%
  mutate(week_day = factor(weekdays(date_time_posix), 
                           #levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
                           levels = c("poniedziałek", "wtorek", "środa", "czwartek", "piątek")
                          ))  

open_queues_preprocessed_data <- data_with_queuers_count %>% filter(liczbaCzynnychStan > 0 & status == 1 )


# Avg nr of served people per day
plot_avg_served_people_per_day <- function() {
  open_queues_preprocessed_data %>% 
    group_by(name, date) %>% 
    summarise(all_served_people = max(served_people)) %>% 
    group_by(name) %>% 
    summarise(`Mean daily nr of served people` = mean(all_served_people)) %>% 
    arrange(desc(`Mean daily nr of served people`)) %>%
    mutate(Office=reorder(name, `Mean daily nr of served people`)) %>% 
    ggplot(aes(x = Office, y = `Mean daily nr of served people`)) +
    geom_col() + coord_flip() + theme_bw()
}
# Avg nr of people in queue per day
plot_avg_queue_len <- function() {
  open_queues_preprocessed_data %>% 
    group_by(name, date) %>% 
    summarise(avg_queue_len = mean(liczbaKlwKolejce)) %>% 
    group_by(name) %>% 
    summarise(`Mean daily queue length` = mean(avg_queue_len)) %>% 
    #summarise(avg_queue_len_per_day = mean(liczbaKlwKolejce)) %>% 
    mutate(Office=reorder(name, `Mean daily queue length`)) %>% 
    ggplot(aes(x = Office, y = `Mean daily queue length`)) +
    geom_col() + coord_flip() + theme_bw()
}


common_week_data <- data_with_queuers_count %>% 
  filter(date_time_posix > as.POSIXct(strptime("2020-01-07", format="%Y-%m-%d", tz="Europe/Warsaw")) &
          date_time_posix < as.POSIXct(strptime("2020-01-20", format="%Y-%m-%d", tz="Europe/Warsaw"))) 
plot_common_week_data <- function() {
  
  
  common_week_data %>% 
    filter(liczbaCzynnychStan > 0) %>%  # filters out days that offices doesn't work
    mutate(`Day of weeek` = week_day) %>% 
    group_by(`Day of weeek`) %>% 
    summarise(`Mean queue length` = mean(liczbaKlwKolejce)) %>% 
    ggplot(aes(x = `Day of weeek`, y = `Mean queue length`)) +
    geom_col() +
    theme_bw() +
    theme( axis.text.x = element_text(angle=45, hjust = 1))
}

plot_served_people_office_queue <- function() {
  jitter_served_people <- data_with_queuers_count %>% 
    filter(liczbaCzynnychStan > 0 & status == 1 ) %>% # filters out days that offices doesn't work
    group_by(name, nazwaGrupy, date) %>% 
    summarise(all_served_people = max(served_people)) %>% 
    group_by(name, nazwaGrupy) %>% 
    summarise(`Mean daily nr of served people` = mean(all_served_people)) %>% 
    mutate(Office = name)
  
  pos = position_jitter(width = 0.5, seed = 1)
  ggplot(jitter_served_people, aes(x = Office, y = `Mean daily nr of served people`, color=name)) +
    geom_point(position = pos ) +
    geom_text_repel(
      aes(label=ifelse(`Mean daily nr of served people` > 250, as.character(nazwaGrupy), '')),
      box.padding   = 0.35, 
      point.padding = 0.5,
      segment.color = 'grey50',
      position = pos ) +
    theme_bw() +
    scale_colour_manual(values = cols25(n=21)) +
    theme(axis.text.x = element_text(angle = 45, 
                                     hjust = 1),
          legend.position = "none")
}

# Avg nr of people in queue per day
plot_queue_len_office_queue <- function() {
  jitter_queue_len <- data_with_queuers_count %>% 
    filter(liczbaCzynnychStan != 0) %>% # filters out days that offices doesn't work
    group_by(name, nazwaGrupy, date) %>% 
    summarise(all_queued_people = mean(liczbaKlwKolejce)) %>% 
    group_by(name, nazwaGrupy) %>% 
    summarise(`Mean queue length` = mean(all_queued_people))
  
  pos = position_jitter(width = 0.5, seed = 2)
  ggplot(jitter_queue_len, aes(x = name, y = `Mean queue length`, color=name)) +
    geom_point(position = pos ) +
    geom_text_repel(
      aes(label=ifelse(`Mean queue length` > 7, as.character(nazwaGrupy), '')),
      box.padding   = 0.35, 
      point.padding = 0.5,
      segment.color = 'grey50',
      position = pos ) +
    theme_bw() +
    scale_colour_manual(values = cols25(n=21)) +
    theme(axis.text.x = element_text(angle = 45, 
                                     hjust = 1),
          legend.position = "none")
}

office_queues_ids <- open_queues_preprocessed_data %>%
  group_by(name, nazwaGrupy, idGrupy, date) %>%
  summarise(n = n(), served = max(served_people), mean_queuers = mean(liczbaKlwKolejce)) %>% 
  arrange(desc(mean_queuers))
queue_efficiency_map <- function() {
  
  pos = position_jitter(width = 0.5, seed = 3)
  office_queues_ids %>% mutate(`Mean queue length` = mean_queuers, `Served people per day` = served, Office = name) %>% 
    ggplot(aes(x = `Served people per day`, y = `Mean queue length`, color = Office)) +
    geom_point() +
    geom_text_repel(
      aes(label=ifelse(mean_queuers > 50, as.character(nazwaGrupy), '')),
      box.padding   = 0.35, 
      point.padding = 0.5,
      segment.color = 'grey50',
      position = pos )+
    theme_bw() +
    scale_colour_manual(values = cols25(n=21))
}


office_queue_data_prepare <- function (data, office_name, queue_id) {
   data %>% 
    filter(idGrupy == queue_id & 
             name == office_name & 
             liczbaCzynnychStan != 0) %>% 
    mutate(Time = time_posix,
           `Served people` = served_people,
           `Queue length` = liczbaKlwKolejce)
}

mokotow_658_queue <- office_queue_data_prepare(common_week_data, "UD_Mokotow_1", "658")
#UD_Praga_Poludnie	Rejestracja pojazdów - składanie wniosków	165	2020-01-08	127	167	29.779528
praga_poludnie_165_queue <- office_queue_data_prepare(common_week_data, "UD_Praga_Poludnie", "165")

wrap_plot_queues <- function(office_data, y_var) {
  office_data %>% ggplot(aes(x=Time, y=!!enquo(y_var), color=date)) +
    geom_point() +
    scale_x_datetime() +
    scale_colour_ordinal() +
    theme_bw() +
    ggtitle(label = office_data[1, "name"], subtitle = office_data[1, "nazwaGrupy"])
}
# Graph showing how nr of served people grows for different time slots and dates
wrap_plot_queues(mokotow_658_queue, 
                 `Served people`)

# Graph showing how nr of queued people changes for different time slots and dates
wrap_plot_queues(mokotow_658_queue, 
                 `Queue length`)

wrap_plot_queues(praga_poludnie_165_queue, 
                 `Served people`)

# Graph showing how nr of queued people changes for different time slots and dates
wrap_plot_queues(praga_poludnie_165_queue, 
                 `Queue length`)

mokotow_10_00 <- data_with_queuers_count %>%
  filter(name == "UD_Mokotow_1" & time == "10:00")

mokotow_10_00 

queues_for_12_00_grp <- data_with_queuers_count %>% filter(time == "12:00" & liczbaCzynnychStan > 0) %>%  
  pivot_longer(cols=c(liczbaCzynnychStan, liczbaKlwKolejce), names_to = "Type", values_to= "Count") %>% 
  mutate(Count = ifelse(Type == "liczbaCzynnychStan", -Count, Count)) %>% 
  group_by(name, nazwaGrupy)

queues_for_12_00_grp %>% filter(name == "UD_Mokotow_1") %>%  
  ggplot(aes(x = nazwaGrupy, y = Count, fill = Type)) +
  geom_bar(stat = "identity", width = .6) +
  coord_flip()




plot_wait_time_per_timeslot <- function(queue_data, office_name, queue_id, date_str) {
  
  mokotow_queue_with_served_number <- queue_data %>% filter(name == office_name)
  
  data_with_slot_nr <- mokotow_queue_with_served_number %>%
    group_by(date, idGrupy ) %>%
    arrange(time) %>%
    mutate(slot_nr=row_number(), next_slot_nr=row_number()+1) %>%
    ungroup()
  
  mokotow_x_with_slot_nr <- data_with_slot_nr %>%  
    filter(idGrupy  == queue_id & date == date_str & liczbaCzynnychStan > 0)
  
  if (all(mokotow_x_with_slot_nr[["liczbaKlwKolejce"]] == 0)) 
    stop("Error no people in the queue!")
  
  if (all(mokotow_x_with_slot_nr$served_people == 0)) 
    stop("Error no people served in the queue!")
  
  mokotow_x_with_slot_time_limits <- mokotow_x_with_slot_nr %>%
    inner_join(data_with_slot_nr, 
               data_with_slot_nr, 
               by=c("next_slot_nr"="slot_nr", "date"="date", "literaGrupy"="literaGrupy")) %>% 
    mutate(slot_visitor_number_start = liczbaKlwKolejce.x + served_people.x,
           slot_visitor_number_end = liczbaKlwKolejce.y + served_people.y) %>% 
    select(time.x, time.y, slot_nr, next_slot_nr, slot_visitor_number_start,
           slot_visitor_number_end, served_people.x, served_people.y, liczbaKlwKolejce.x,
           liczbaKlwKolejce.y, time_posix.x, time_posix.y, liczbaCzynnychStan.x) %>% 
    rename(slot_start = time.x,
           slot_end = time.y,
           slot = slot_nr,
           next_slot = next_slot_nr,
           start_last_queuer_nr = slot_visitor_number_start,
           end_last_queuer_nr = slot_visitor_number_end,
           start_served_nr = served_people.x, 
           end_served_nr = served_people.y,
           start_queue_length = liczbaKlwKolejce.x, 
           end_queue_length = liczbaKlwKolejce.y,
           start_time = time_posix.x,
           end_time = time_posix.y, 
           active_lines_count = liczbaCzynnychStan.x
           )
  #browser()
  
  mokotow_x_self_cross_joined <- mokotow_x_with_slot_time_limits %>%
    mutate(dummy = TRUE) %>%
    inner_join(mokotow_x_with_slot_time_limits %>% mutate(dummy = TRUE),
               by = c("dummy" = "dummy")) %>% 
    select(-dummy) 
  
  
  # .x
  #        end_last_queuer_nr
  #came    |            |          |
  #     xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
  #served|    |                       |      
  #           end_served_nr            
  
  
  # .y
  #
  #came    |            |          |
  #     xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
  #served|    |                       |      
  
  # y - serverd
  # x - queued
  
  # any person that enqueued in given slot.x match with slot.y that served any of people enqueued in that slot
  # start_qued_person < serving_person_start & serving_person_end < end_qued_person |
  # serving_person_start < end_qued_person < serving_person_end | serving_person_start < start_qued_person < serving_person_end
  
  mokotow_x_paired_served_enqueued_slots <- mokotow_x_self_cross_joined %>% filter(
    start_last_queuer_nr.x < start_served_nr.y & end_served_nr.y < end_last_queuer_nr.x |
      start_served_nr.y <= end_last_queuer_nr.x & end_last_queuer_nr.x < end_served_nr.y |
      start_served_nr.y < start_last_queuer_nr.x & start_last_queuer_nr.x <= end_served_nr.y
    ) 
  
  mokotow_x_slots_wait_time <- mokotow_x_paired_served_enqueued_slots %>%
    filter(start_last_queuer_nr.x < end_last_queuer_nr.x) %>% 
    mutate(wait_time = as_hms(pmax(start_time.y + (end_time.y - start_time.y)/2 - (start_time.x + (end_time.x - start_time.x)/2), 0)),
           time_slot = as_hms(start_time.x + (end_time.x - start_time.x)/2),
           slot_start = as_hms(start_time.x),
           width = end_time.x - start_time.x)
    
  
  mokotow_x_avg_wait_time <- mokotow_x_slots_wait_time %>% 
    group_by(time_slot, width) %>%
    summarise(
      avg_wait_time=as_hms(mean(wait_time))
    )
  
  slot_even = as.logical(1:nrow(mokotow_x_avg_wait_time)%%2)
  
  # Plot showing how long on average people had to wait until getting served
  mokotow_x_avg_wait_time %>% 
    mutate(`Time` = time_slot,
           `Mean waiting time` = avg_wait_time,
           `Time slot [sec]` = as.numeric(width, "secs")) %>% 
    ggplot(aes(x = `Time`, 
               y = `Mean waiting time`,
               width =`Time slot [sec]`, #as.numeric(mokotow_x_avg_wait_time$width, "secs"),
               fill=`Time slot [sec]`)) + #as.numeric(mokotow_x_avg_wait_time$width, "secs") )) +
    geom_bar(stat = "identity" ) +
    #geom_point() +
    ggtitle(label = paste(mokotow_x_with_slot_nr[[1, "name"]], mokotow_x_with_slot_nr[[1, "date"]]),
            subtitle = mokotow_x_with_slot_nr[[1, "nazwaGrupy"]]) +
    theme_bw()
}




common_office_queues_ids <- common_week_data %>%
  group_by(name, nazwaGrupy, idGrupy, date) %>%
  summarise(n = n(), served = max(served_people), mean_queuers = mean(liczbaKlwKolejce)) %>%
  arrange(desc(mean_queuers))

#View(common_office_queues_ids)
#View(office_queues_ids %>% arrange(desc(mean_queuers)))



```


# KolejkeR

### API for Otwarte dane kolejkowe po warszawsku (Open Warsaw Queue Data).

## 1. Installation

You can install the development version from
[GitHub](https://github.com/) with:

``` {r installation}
# install.packages("devtools")
#devtools::install_github("HaZdula/kolejkeR")
library(kolejkeR)
```

## 2. Get access to data regarding queues in municipal offices in Warsaw

The city hall of Warsaw made a variety of data public [here](https://api.um.warszawa.pl). Everyone is welcome to access the data via API, using JSON file format. 

One of the datasets contains the information about queues in municipal offices in Warsaw, ie. expected times, open counters and amount of people waiting. 

`kolejkeR` is an R wrapper around the API, developed as a project for Advanced R classes at the Warsaw University of Technology. It's suitable both for further data processing and analysis, as well as for quick acquirement of information helpful for everyday life in Warsaw.

## 3. The main function

The bulwark of job is done by function `get_raw_data(office_name, queue_name)`. This function is designed for data processing and analysis.

The parameters are `character`. You can get their possible values using *getter* functions described in next section.

The function returns a `data.frame` with following columns:

 *  status - either 0 (queue is not operating) or 1 (queue is operating).
 *  czasObslugi - expected time of waiting in queue, in minutes. 
 *  lp - ordinal number.
 *  idGrupy - ID of a queue from `nazwaGrupy`.
 *  liczbaCzynnychStan - amount of opened counters. 
 *  nazwaGrupy - a name of a queue.
 *  literaGrupy - a single letter symbolizing a queue name from `nazwaGrupy`. 
 *  liczbaKlwKolejce - amount of people in queue.
 *  aktualnyNumer - current ticket number.

## 4. Getters

2 functions return acceptable values, that can be passed either to `get_raw_data()` or to functions described below.

* `get_available_offices()` returns names of municipal offices in Warsaw, accepted as `office_name` arguments in other functions.

* `get_available_queues(office_name)` returns names of queues in the chosen municipal office in Warsaw, accepted as `queue_name` arguments in other functions.

## 5. Getting specific data directly

These functions, vectorized over `queue_name`, are designed to get exact data quickly in form of a `character` vector of the same length as `queue_name`. These are:

* `get_waiting_time(office_name, queue_name)`

* `get_open_counters(office_name, queue_name)`

* `get_current_ticket_number(office_name, queue_name)`

* `get_number_of_people(office_name, queue_name)`.

## 6. Getting specific data verbosely

These functions, vectorized over `queue_name`, are designed to get exact data quickly in form of a user-friendly statement. Each of them returns a `character` vector of the same length as `queue_name`, with each element in a form of a sentence , depending the question called. These are:

* `get_waiting_time(office_name, queue_name, language = 'en')`

* `get_open_counters(office_name, queue_name, language = 'en')`

* `get_current_ticket_number(office_name, queue_name, language = 'en')`

* `get_number_of_people(office_name, queue_name, language = 'en')`.

The `language` variable regards the desired language of the output. Currently the only supported languages are English and Polish.

## 7. Example

Let's assume, that you would like to file an application for 500+ programme in municipal office in Bielany in Warsaw. You wish to find out the average time you need to wait and the amount of people already waiting.

``` {r}
library(kolejkeR)
library(stringi)

offices <- get_available_offices()

# Find out the exact name of the office
Bielany_office <- offices[stri_detect_fixed(offices, "Bielany")][1]
queues <- get_available_queues(Bielany_office)

# Find the exact name of the queue
queue500 <- queues[stri_detect_fixed(queues, "500+")][1]

print(get_waiting_time_verbose(Bielany_office, queue500, language = 'pl'))
print(get_number_of_people_verbose(Bielany_office, queue500, language = 'pl'))

```
Now you can go to the municipal office with a big smile on your face.

## 8. Dataset

Package function were used to gather data about queues in Warsaw's government offices. Data was collected on the span of couple of weeks from December 28 to January 22.
Warsaw's office API was called repeatedly every 5 minutes and results appended to csv file. KolejkeR package includes this dataset in a form of a data frame with 830172 rows and 12 variables. Uncompressed csv file with dataset takes 84.3 MB of memory.

## 9. Dataset analysis

Script included to vignettes folder in the package stores some data analysis of the dataset. Libraries used for analysis and plotting graph are added to suggested dependencies. Results of the analysis are presented below.


``` {r, echo = FALSE, eval = TRUE}
plot_avg_served_people_per_day()
```


Above plot shows how many people are served in Warsaw offices every day on average.


``` {r, echo = FALSE, eval = TRUE}
plot_avg_queue_len()
```


This graph shows the length of the queues in Warsaw's offices. Office with longest queues on average has very small number of dedicated queues and is the second office in the same district.  


``` {r, echo = FALSE, eval = TRUE}
plot_common_week_data()
```


Average queue lengths for different weekdays show that on average people tend to go to offices slightly more often on the beginning of the week. That maybe because some offices work longer at the begging of the week and longer working hours draw more people.


``` {r, echo = FALSE, eval = TRUE}
plot_served_people_office_queue()
```


Number of served people every day depends strongly on the queue objective and office.


``` {r, echo = FALSE, eval = TRUE}
plot_queue_len_office_queue()
```


On average longest queues are for vehicle registrations.


``` {r, echo = FALSE, eval = TRUE}
queue_efficiency_map()
```


Map queue objective description on mean queue length and average number of served people per day shows efficiency of given serving booths. Despite long queues relatively small number of people is served for vehicle registration and other labelled queues. This indicates processes that take long time to achieve in the offices. 


``` {r, echo = FALSE, eval = TRUE}
# Graph showing how nr of served people grows for different time slots and dates
wrap_plot_queues(mokotow_658_queue, 
                 `Served people`)
```


Growth of served people over time for example queue for couple of different dates shows linearity.


``` {r, echo = FALSE, eval = TRUE}

# Graph showing how nr of queued people changes for different time slots and dates
wrap_plot_queues(mokotow_658_queue, 
                 `Queue length`)

```


Changes of queue length for sample queue show more interesting relationships. Queues for this queue were longer in afternoon hours. Monday (2020-01-13) was the most popular weekday. 


``` {r, echo = FALSE, eval = TRUE}
wrap_plot_queues(praga_poludnie_165_queue, 
                 `Served people`)

```


The same graph for vehicle registration show that this process takes much longer to serve single person in the queue.


``` {r, echo = FALSE, eval = TRUE}
# Graph showing how nr of queued people changes for different time slots and dates
wrap_plot_queues(praga_poludnie_165_queue, 
                 `Queue length`)
```


For the more popular queue people tend to come mostly before noon.


``` {r, echo = FALSE, eval = TRUE}
plot_wait_time_per_timeslot(data_with_queuers_count, "UD_Bemowo", "551", "2020-01-13")
```


``` {r, echo = FALSE, eval = TRUE}
plot_wait_time_per_timeslot(data_with_queuers_count, "UD_Bielany", "171", "2020-01-13")
```


``` {r, echo = FALSE, eval = TRUE}
plot_wait_time_per_timeslot(data_with_queuers_count, "UD_Ochota", "599", "2020-12-30")
```


``` {r, echo = FALSE, eval = TRUE}
plot_wait_time_per_timeslot(data_with_queuers_count, "UD_Mokotow_1", "658", "2020-12-30")
```


Plots showing distribution of waiting time for people that come in different hours show that waiting time for serving queuers is non-obvious and traces individual characteristics for different queues.
This plots were created with assumption that ticket nr indicates how many people were served for given moment and no one is leaving queue before getting ticket and being served. 
Blank columns indicate that this assumption not always holds. Especially for queues were many serving booths were active ticket nr tends to grow slower than queue length reduction.

